---
title: "Geomorphology_Ellerbe"
author: "KWhitmore"
date: "08/08/2024"
output: html_document
---

#analysis brainstorm
1. Can I model discharge and constitute concentration/discharge assuming constant constitute loading to Ellerbee creek?
2. Can I model but based on Josephs data (or other model)?
3. And then I can compare 1 and 2 with the actual data (though I didn't include discharge, so that will have to be modeled --- I do have a couple USGS stations. AND Tommy maybe??)


4. SWAT?



urine water
lawn fertilizer (though that's usually ammonium)

look at WRR 2022 paper. isotopic sourcing
HP 
look at graphs look at predictors

Johnny ecoli

kashal synoptic sampling  - frontiers 2023

 > cummulative upstream land use
 > what's the delta 7
 
pick points of interest`    


J drive: where the data is at
DEM might be on the j drive


email joseph for j drive exploration

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(lubridate)
library(plotly)
library(dplyr)
library(tidyr)
library(zoo)
#library(plot3D)
library(splines)
library(Ecdat)
library(dplyr)
library(sp)
library(geosphere)
library(ggmap)

```

## Build the  data frame for EllerbeCreek Tracks


```{r build df}

#2021-10-20 18:24 - 19:00
Geomorph <- read.csv(here::here("FallData/GPS/Ellerbe_Synop_2021-10-20.csv"), skip = 42)[,3:6]

#convert time in UTC to local time
Geomorph$DateTime <- gsub("T"," ", Geomorph$time)
Geomorph$DateTime <- gsub("Z","", Geomorph$DateTime)
Geomorph$DateTime <- as.POSIXct(Geomorph$DateTime, format = "%Y-%m-%d %H:%M:%S", tz = "UTC" )
Geomorph$DateTime <-Geomorph$DateTime - 4*60*60 
#round to nearest half second
Geomorph$DateTime <- round_date(Geomorph$DateTime, unit = "15 second")
Geomorph_1 <- subset(Geomorph,
      DateTime > as.POSIXct('2021-10-20 18:30:00', tz="UTC") & DateTime < as.POSIXct('2021-10-20 19:00:00', tz="UTC"))


#2021-10-21 08:30 - 17:50
Geomorph <- read.csv(here::here("FallData/GPS/Ellerbe_Synop_2021-10-21.csv"), skip = 42)[,3:6]
#convert time in UTC to local time
Geomorph$DateTime <- gsub("T"," ", Geomorph$time)
Geomorph$DateTime <- gsub("Z","", Geomorph$DateTime)

Geomorph$DateTime <- as.POSIXct(Geomorph$DateTime, format = "%Y-%m-%d %H:%M:%S", tz = "UTC" )
Geomorph$DateTime <-Geomorph$DateTime - 4*60*60 
#round to nearest half second
Geomorph$DateTime <- round_date(Geomorph$DateTime, unit = "15 second")

Geomorph_2 <- subset(Geomorph,
      DateTime > as.POSIXct('2021-10-21 08:29:00', tz="UTC") & DateTime < as.POSIXct('2021-10-21 17:50:00', tz="UTC"))

#2021-10-22 09:10 - 14:30
Geomorph <- read.csv(here::here("FallData/GPS/Ellerbe_Synop_2021-10-22.csv"), skip = 42)[,3:6]
#convert time in UTC to local time
Geomorph$DateTime <- gsub("T"," ", Geomorph$time)
Geomorph$DateTime <- gsub("Z","", Geomorph$DateTime)

Geomorph$DateTime <- as.POSIXct(Geomorph$DateTime, format = "%Y-%m-%d %H:%M:%S", tz = "UTC" )
Geomorph$DateTime <-Geomorph$DateTime - 4*60*60 
#round to nearest half second
Geomorph$DateTime <- round_date(Geomorph$DateTime, unit = "15 second")

Geomorph_3 <- subset(Geomorph,
      DateTime > as.POSIXct('2021-10-22 09:10:00', tz="UTC") & DateTime < as.POSIXct('2021-10-22 14:30:00', tz="UTC"))

Geomorph <- rbind(Geomorph_1,Geomorph_2,Geomorph_3)
distance_df <- rbind(Geomorph_3,Geomorph_2)
rm(Geomorph_1,Geomorph_2,Geomorph_3)
```

#read in water samples

```{r add synop to df}
WaterSamples <- read.csv(here::here("FallData/WaterSamples/UNC_11-11-2021.csv"))

colnames(WaterSamples) <- c("SampleName","Cl_mgL","SO4_mgL","Br_mgL","NO3_mgL", "DOC_mgL","TDN_mgL")

WaterSamples$SampleNo <-  as.numeric(sub("(_).*","", WaterSamples$SampleName))
WaterSamples$Date <- as.Date(gsub(".*_(.+)_.*", "\\1", WaterSamples$SampleName), format = "%m-%d-%y")
WaterSamples$Time <- sub(".*_", "\\1", WaterSamples$SampleName)

WaterSamples$DateTime <- as.POSIXct(paste(WaterSamples$Date, WaterSamples$Time), format="%Y-%m-%d %H:%M", tz="UTC")
WaterSamples$DateTime <- round_date(WaterSamples$DateTime, unit = "1 minute")

```

#plot 

Lat lon for water treatment out flow:
36.031767, -78.862275


```{r look at all synop data}
df <- full_join(WaterSamples,Geomorph, by="DateTime") 

#TrackMap <- qmplot(lon, lat, data = df#%>%drop_na(DOC_mgL)
#                   , zoom = 13, #color = grey,
#                   maptype = "stamen_terrain"
#                   ) 
TrackMap +
  geom_point(aes(color = DOC_mgL), data = df%>%drop_na(DOC_mgL), size = 3) +
  scale_color_gradient(low="blue", high="red")

TrackMap +
  geom_point(aes(color = log10(NO3_mgL)), data = df%>%drop_na(DOC_mgL), size = 3) +
  scale_color_gradient(low="blue", high="red") 

TrackMap +
  geom_point(aes(color = log10(Cl_mgL)), data = df%>%drop_na(DOC_mgL), size = 3) +
  scale_color_gradient(low="blue", high="red") 

TrackMap +
  geom_point(aes(color = log10(Br_mgL)), data = df%>%drop_na(DOC_mgL), size = 3) +
  scale_color_gradient(low="blue", high="red") 
TrackMap +
  geom_point(aes(color = log10(SO4_mgL)), data = df%>%drop_na(DOC_mgL), size = 3) +
  scale_color_gradient(low="blue", high="red") 


```

#calibrate data

```{r calibrated data}

Mainstem_predict <- read.csv(here::here("FallData/KW_data_prelimcal2024-06-18.csv"))%>%rename(DateTime = TIMESTAMP)

Mainstem_predict$DateTime <- as.POSIXct(Mainstem_predict$DateTime, format = "%Y-%m-%d %H:%M:%S", tz = "UTC" )


df_predict <- full_join(df,dropped_nas, by = "DateTime")

#df_predict_dropped_nas <- df_predict |>
  #drop_na(NO3.N) |>
  #mutate(log10_NO3N = log10(NO3.N))

TrackMap <- qmplot(lon, lat, data = df_predict%>%drop_na(DOC), zoom = 13, maptype = "stamen_terrain") 

TrackMap +
  geom_point(aes(color = log10(NO3.N)), data = df_predict %>%drop_na(DOC), size = 2) +
  scale_color_gradient(low="blue", high="red") +
  labs(color = "log10(NO3 as N \n (mg/L))") 

TrackMap +
  geom_point(aes(color = log10(DOC)), data = df_predict%>%drop_na(DOC), size = 2) +
  scale_color_gradient(low="blue", high="red") +
  labs(color = "log10(DOC \n (mg/L))") 


#write.csv(df_predict_dropped_nas,here::here("calibrated_forJesse_dropped_nas2.csv"))
```
#zoom in

zoom in to area upstream of water treatment plant so we can see some variation in DOC and NO3 concentrations

```{r zoom in}

data_upstream <- subset(df_predict, DateTime > as.POSIXct('2021-10-21 13:45:00', tz="UTC"))


#TrackMap <- qmplot(lon, lat, data = data_upstream%>%drop_na(DOC), zoom = 13, 
 #                  maptype = "stamen_terrain") 

TrackMap +
  geom_point(aes(color = NO3.N), data = data_upstream %>%drop_na(DOC), size = 2) +
  scale_color_gradient(low="blue", high="red") +
  labs(color = "NO3 as N \n (mg/L)") 

TrackMap +
  geom_point(aes(color = DOC), data = data_upstream %>%drop_na(DOC), size = 2) +
  scale_color_gradient(low="blue", high="red") +
  labs(color = "DOC \n (mg/L)") 

```

#include distance
we will remove s. Ellerbe brance for now

```{r  include distance}

distance_df <- subset(df_predict,DateTime > as.POSIXct('2021-10-21 08:29:00', tz="UTC"))
#order by decreasing time
distance_df <- distance_df[order(-as.numeric(distance_df$DateTime)),]

###loop da loop

distance_df <- distance_df%>%rename(lon_fit=lon)%>%rename(lat_fit=lat)%>%drop_na(lat_fit)

distance_df$lon_dist <- NA
distance_df$lat_dist <- NA
i <- 2
for(i in 1:nrow(distance_df)) {
  if (i == 1) {

      distance_df$dist <- 0
      
    } else {
      distance_df[i,"lon_dist"] <- distance_df[i-1,"lon_fit"]
      distance_df[i,"lat_dist"] <- distance_df[i-1,"lat_fit"]
      distance_df[i,"dist_diff"] <- distHaversine(
        c(distance_df[i,"lon_fit"], distance_df[i,"lat_fit"]), 
        c(distance_df[i,"lon_dist"], distance_df[i,"lat_dist"]))
      distance_df[i,"dist"] <- distance_df[i,"dist_diff"] + distance_df[i-1,"dist"]
      }
}

```

# plot by distance

```{r  include distance}

p1 <- ggplot(data=distance_df,aes(x=dist,y=log(NO3.N))) + geom_point(size=1)+
  scale_color_gradient(low="blue", high="red")
p1

p2 <- ggplot(data=distance_df ,aes(x=dist,y=log(DOC))) + geom_point(size=1)+
  scale_color_gradient(low="blue", high="red")
p2

p3 <- ggplot(data=distance_df ,aes(x=dist,y=log(TDN))) + geom_point(size=1)+
  scale_color_gradient(low="blue", high="red")
p3

p5 <- ggplot(data=distance_df,aes(x=log(NO3.N),y=log(DOC),color=dist)) + geom_point(size=1)
p5

p6 <- ggplot(data=distance_df,aes(x=log(TDN),y=log(DOC),color=dist)) + geom_point(size=1)
p6
```
